---
title: "Sentida"
author: "Adela Sobotkova"
date: "2025-3-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This document expands the IPCC sentiment analysis with a Danish implementation of Sentida package. It requires the same packages as IPCC plus the Danish-specific resources. In short, run W11 first, then this script. 

## Load the Sentida package

```{r}
if(!require("devtools")) install.packages("devtools")

devtools::install_github("Guscode/Sentida")

library(Sentida)
```

Now that we have the toolkit, locate a suitable Danish dataset. 
Let's see about the mediastream subset


## Load Danish stopwords

See `?stop_words` and `View(stop_words)`to look at documentation for stop words lexicons.
For Danish stopwords, see here:  
https://raw.githubusercontent.com/stopwords-iso/stopwords-da/refs/heads/master/stopwords-da.txt

```{r}
dk_stopwords <- read.csv("https://raw.githubusercontent.com/stopwords-iso/stopwords-da/refs/heads/master/stopwords-da.txt", header = FALSE)

new_stopwords <- data_frame(word = c("þÿ", "paa", "a", "be", "cf", "r", "t", "saa", "1", "kr", "m", "f", "al", "maa", "al", "ogsaa", "uden", "mød", "v", "r", "e", "n", "ca", "pa", "re", "reg"))

dk_stopwords <- dk_stopwords %>% 
  rename(word = V1) %>% 
  rbind(new_stopwords)
```


```{r}

# random cross checking 
sentida("Jeg er blevet mere eventyrlysten.", output = "total")

sentida("Udvidet horisonter Ting man fandt ekstreme og unormale er med tiden blevet mere normaliseret", output = "mean") 


# Analyzing sentiment for each word separately
sentida("Udvidet", output = "mean")        # Sentiment for "Udvidet"
sentida("horisonter", output = "mean")     # Sentiment for "horisonter"
sentida("Ting", output = "mean")           # Sentiment for "Ting"
sentida("man", output = "mean")            # Sentiment for "man"
sentida("fandt", output = "mean")          # Sentiment for "fandt"
sentida("ekstreme", output = "mean")       # Sentiment for "ekstreme"
sentida("og", output = "mean")             # Sentiment for "og"
sentida("unormale", output = "mean")       # Sentiment for "unormale"
sentida("er", output = "mean")             # Sentiment for "er"
sentida("med", output = "mean")            # Sentiment for "med"
sentida("tiden", output = "mean")          # Sentiment for "tiden"
sentida("blevet", output = "mean")         # Sentiment for "blevet"
sentida("mere", output = "mean")           # Sentiment for "mere"
sentida("normaliseret", output = "mean")   # Sentiment for "normaliseret"

sentida(iconv("Udvidet horisonter Ting man før fandt ekstreme og unormale er med tiden blevet mere normaliseret","latin1", "UTF-8") , output = "total") 

```
